# Spatial sequencing based technology, Visium Part I

Joselyn Cristina Chávez Fuentes

August 5th 2024

## Introduction to spatial dataset

10X genomics recently launched a new platform to obtain spatial expression data using a Visium Spatial Gene Expression slide.

The Visium brain data to run this tutorial can be found [here](https://support.10xgenomics.com/spatial-gene-expression/datasets/1.1.0/V1_Adult_Mouse_Brain)

## Create the Giotto object

```{r, eval=FALSE}
library(Giotto)

## Set instructions
results_folder <- "results/"

my_python_path <- NULL

instrs <- createGiottoInstructions(
    save_dir = results_folder,
    save_plot = TRUE,
    show_plot = FALSE,
    python_path = my_python_path
)

## Provide path to visium folder
data_path <- "data/"

## Create object directly from visium folder
visium_brain <- createGiottoVisiumObject(
    visium_dir = data_path,
    expr_data = "raw",
    png_name = "tissue_lowres_image.png",
    gene_column_index = 2,
    instructions = instrs
)

## show plot
spatPlot2D(
    gobject = visium_brain,
    cell_color = "in_tissue",
    point_size = 2,
    cell_color_code = c("0" = "lightgrey", "1" = "blue"),
    show_image = TRUE,
    image_name = "image"
)
```


## Subset on spots that were covered by tissue

```{r, eval=FALSE}
metadata <- pDataDT(visium_brain)
in_tissue_barcodes <- metadata[in_tissue == 1]$cell_ID
visium_brain <- subsetGiotto(visium_brain,
    cell_ids = in_tissue_barcodes
)
```


## Quality control

- Statistics

## Filtering

```{r, eval=FALSE}
visium_brain <- filterGiotto(
    gobject = visium_brain,
    expression_threshold = 1,
    feat_det_in_min_cells = 50,
    min_det_feats_per_cell = 1000,
    expression_values = c("raw"),
    verbose = TRUE
)
```


## Normalization


```{r, eval=FALSE}
visium_brain <- normalizeGiotto(
    gobject = visium_brain,
    scalefactor = 6000,
    verbose = TRUE
)
```


## Feature selection

text 
 
 
### Highly Variable Features:

- loess regression
- binned
- pearson residuals

### Spatial variable genes

text 

## Dimension Reduction

- PCA
- UMAP/t-SNE
- Visualizations

## Clustering

text 

### Non-spatial

- k-means
- Hierarchical clustering
- Leiden/Louvain

### Spatial

- Spatial variable genes
- Spatial co-expression modules



