# Creating a Giotto object

Jiaji George Chen

August 5th 2024


```{r, eval=FALSE}
save_dir <- "~/Documents/GitHub/giotto_workshop_2024/img/01_session4"
```

## Overview

Giotto has representations for both aggregated (cell by count) + xy(z) 
information and subcellular polygon or mask information and transcript xy(z)
or image information. A Giotto object can be set up from either of the two
above sets of information.

## From matrix + locations

```
createGiottoObject()
```

## From subcellular raw data (transcripts or images) + polygons

```
createGiottoObjectSubcellular()
```

The above two methods accept data, converts them into Giotto's compatible
formats, and then assembles the final giotto object.

If desired you can also assemble a giotto object piece-wise after creating
the Giotto subobjects manually.

## From piece-wise

```
g <- giotto()
g <- setGiotto(g, ??)
```

## Using convenience functions for popular technologies (Vizgen, Xenium, CosMx, …)

There are also several convenience functions we provide for loading in data
from popular platforms. Many of these will be touched on later during other
sessions
```
createGiottoVisiumObject()
createGiottoVisiumHDObject()
createGiottoXeniumObject()
createGiottoCosMxObject()
createGiottoMerscopeObject()
```

## Spatial plots

Giotto has several spatial plotting functions.
At the lowest level, you directly call `plot()` on several subobjects in order
to see what they look like, particularly the ones containing spatial info.

```{r, eval=FALSE}
gpoints <- GiottoData::loadSubObjectMini("giottoPoints")
gpoly <- GiottoData::loadSubObjectMini("giottoPolygon")
spatlocs <- GiottoData::loadSubObjectMini("spatLocsObj")
spatnet <- GiottoData::loadSubObjectMini("spatialNetworkObj")
pca <- GiottoData::loadSubObjectMini("dimObj")
plot(pca, dims = c(3,10))
```

## Subsetting

-   Based on IDs
-   Based on locations
-   Visualizations



## Mini objects & GiottoData

Giotto makes available several mini objects to allow devs and users to work
with easily loadable Giotto objects. These are small subsets of a larger
dataset that often contain some worked through analyses and are fully functional.
```{r, eval=FALSE}
pak::pak("drieslab/GiottoData")
```


