# Visium Part II

Joselyn Cristina Chávez Fuentes

August 6th 2024

## Differential expression


- text

## Enrichment & Deconvolution

- PAGE/Rank
- SpatialDWLS
- Visualizations

## Spatial expression patterns

- Spatial variable genes

Rank binarization

```{r, eval=FALSE}
ranktest = binSpect(visium_brain, 
                    bin_method = 'rank',
                    calc_hub = TRUE, 
                    hub_min_int = 5,
                    spatial_network_name = 'spatial_network')
```

Visualize

```{r, eval=FALSE}
spatFeatPlot2D(visium_brain, 
               expression_values = 'scaled',
               feats = ranktest$feats[1:6], 
               cow_n_col = 2, 
               point_size = 1.5)
```

- Spatial co-expression modules

Cluster the top 500 spatial genes into 20 clusters

```{r, eval=FALSE}
ext_spatial_genes = ranktest[1:1500,]$feats
```

Use existing detectSpatialCorGenes function to calculate pairwise distances between genes (but set network_smoothing=0 to use default clustering)

```{r, eval=FALSE}
spat_cor_netw_DT = detectSpatialCorFeats(visium_brain,
                                         method = 'network',
                                         spatial_network_name = 'spatial_network',
                                         subset_feats = ext_spatial_genes)
```

Identify most similar spatially correlated genes for one gene

```{r, eval=FALSE}
top10_genes = showSpatialCorFeats(spat_cor_netw_DT, 
                                  feats = 'Ptprn', 
                                  show_top_feats = 10)
```

Visualize

```{r, eval=FALSE}
spatFeatPlot2D(visium_brain, 
               expression_values = 'scaled',
               feats = top10_genes$variable[1:4], 
               point_size = 3)
```


Cluster spatial genes

```{r, eval=FALSE}
spat_cor_netw_DT = clusterSpatialCorFeats(spat_cor_netw_DT, 
                                          name = 'spat_netw_clus', 
                                          k = 20)
```

Visualize clusters

```{r, eval=FALSE}
heatmSpatialCorFeats(visium_brain,
                     spatCorObject = spat_cor_netw_DT,
                     use_clus_name = 'spat_netw_clus',
                     heatmap_legend_param = list(title = NULL),
                     save_param = list(base_height = 6, base_width = 8, units = 'cm'))
```

Rank spatial correlated clusters and show genes for selected clusters

```{r, eval=FALSE}
netw_ranks = rankSpatialCorGroups(visium_brain,
                                  spatCorObject = spat_cor_netw_DT, 
                                  use_clus_name = 'spat_netw_clus',
                                  save_param = list(base_height = 3, base_width = 5))

top_netw_spat_cluster = showSpatialCorFeats(spat_cor_netw_DT, 
                                            use_clus_name = 'spat_netw_clus',
                                            selected_clusters = 6, 
                                            show_top_feats = 1)
```

Create metagene enrichment score for clusters

```{r, eval=FALSE}
cluster_genes_DT = showSpatialCorFeats(spat_cor_netw_DT, 
                                       use_clus_name = 'spat_netw_clus', 
                                       show_top_feats = 1)

cluster_genes = cluster_genes_DT$clus 
names(cluster_genes) = cluster_genes_DT$feat_ID

visium_brain <- createMetafeats(visium_brain, 
                                feat_clusters = cluster_genes, 
                                name = 'cluster_metagene')

spatCellPlot(visium_brain,
             spat_enr_names = 'cluster_metagene',
             cell_annotation_values = netw_ranks$clusters,
             point_size = 1, 
             cow_n_col = 5, 
             save_param = list(base_width = 15))
```

- Spatially informed clusters

Get the top 30 genes per spatial co-expression cluster

```{r, eval=FALSE}
table(spat_cor_netw_DT$cor_clusters$spat_netw_clus)

coexpr_dt = data.table::data.table(genes = names(spat_cor_netw_DT$cor_clusters$spat_netw_clus),
                                   cluster = spat_cor_netw_DT$cor_clusters$spat_netw_clus)

data.table::setorder(coexpr_dt, cluster)
top30_coexpr_dt = coexpr_dt[, head(.SD, 30) , by = cluster]
my_spatial_genes = top30_coexpr_dt$genes
```

Re-calculate the clustering

```{r, eval=FALSE}
visium_brain <- runPCA(gobject = visium_brain,
                       feats_to_use = my_spatial_genes,
                       name = 'custom_pca')

visium_brain <- runUMAP(visium_brain, 
                        dim_reduction_name = 'custom_pca', 
                        dimensions_to_use = 1:20,
                        name = 'custom_umap')

visium_brain <- createNearestNetwork(gobject = visium_brain,
                                     dim_reduction_name = 'custom_pca',
                                     dimensions_to_use = 1:20, 
                                     k = 5,
                                     name = 'custom_NN')

visium_brain <- doLeidenCluster(gobject = visium_brain, 
                                network_name = 'custom_NN',
                                resolution = 0.15, 
                                n_iterations = 1000,
                                name = 'custom_leiden')

cell_meta = pDataDT(visium_brain)
cell_clusters = unique(cell_meta$custom_leiden)

selected_colors = getDistinctColors(length(cell_clusters))
names(selected_colors) = cell_clusters
```

Visualize

```{r, eval=FALSE}
spatPlot2D(visium_brain, 
           cell_color = 'custom_leiden', 
           cell_color_code = selected_colors, 
           coord_fix_ratio = 1)
```

```{r, eval=FALSE}
plotUMAP(gobject = visium_brain, 
         cell_color = 'custom_leiden', 
         cell_color_code = selected_colors, 
         point_size = 1.5)
```

- Spatial domains HMRF

Do HMRF with different betas on top 30 genes per spatial co-expression module

```{r, eval=FALSE}
hmrf_folder = paste0(results_folder,'/','11_HMRF/')
if(!file.exists(hmrf_folder)) dir.create(hmrf_folder, recursive = T)

HMRF_spatial_genes = doHMRF(gobject = visium_brain,
                            expression_values = 'scaled',
                            spatial_genes = my_spatial_genes, 
                            k = 20,
                            spatial_network_name="spatial_network",
                            betas = c(0, 10, 5),
                            output_folder = paste0(hmrf_folder, '/', 'Spatial_genes/SG_topgenes_k20_scaled'))

visium_brain <- addHMRF(gobject = visium_brain, 
                        HMRFoutput = HMRF_spatial_genes,
                        k = 20, 
                        betas_to_add = c(0, 10, 20, 30, 40),
                        hmrf_name = 'HMRF')

spatPlot2D(gobject = visium_brain, 
           cell_color = 'HMRF_k20_b.40')
```


## Interactive tools

text
