# Interoperability with isolated tools

Wen Wang

August 7th 2024


## Spatial niche trajectory analysis

### Prepare

#### Dataset download
```{bash}
# raw data
# giotto
# ONTraC input
# NT score
```


#### Installation ONTraC
```{bash}
source ~/.bash_profile
conda create -y -n ONTraC python=3.11
conda activate ONTraC
pip install ONTraC[analysis]
```

### Running
#### Processing with Giotto
use the given cell type directly

handling multiple sample (Jeff should present this topic on the 2nd day)
```{r}
library(Giotto)

# Load dataset
giotto_1 = createGiottoObject(expression = 'subset_mouse1_slice221_exp.txt',
                                spatial_locs = 'subset_mouse1_slice221_pos.csv')
data1_meta <- read.csv(file = 'subset_mouse1_slice221_dataset.csv')
data1_meta <- data1_meta[c("Cell_Type")]  # we use the given cell type annotation here, you can make your own annotation in process steps
rownames(data1_meta) <- giotto_1@cell_metadata$cell$rna$cell_ID
giotto_1 <- addCellMetadata(giotto_1, new_metadata = data1_meta)


giotto_2 = createGiottoObject(expression = 'subset_mouse2_slice201_exp.txt',
                              spatial_locs = 'subset_mouse2_slice201_pos.csv')
data2_meta <- read.csv(file = 'subset_mouse2_slice201_dataset.csv')
data2_meta <- data2_meta[c("Cell_Type")]  # we use the given cell type annotation here, you can make your own annotation in process ste
rownames(data2_meta) <- giotto_2@cell_metadata$cell$rna$cell_ID
giotto_2 <- addCellMetadata(giotto_2, new_metadata = data2_meta)

# Join giotto_objs
giotto_obj = joinGiottoObjects(gobject_list = list(giotto_1, giotto_2),
                              gobject_names = c('mouse1_slice221', 'mouse2_slice201'),  # name for each samples
                              join_method = 'z_stack')

# Processing the data set using Giotto
# skipped here

# Generate ONTraC input file
pos_df = giotto_obj@spatial_locs$cell[[1]]@coordinates
meta_df = giotto_obj@cell_metadata$cell$rna@metaDT

output_df = merge(x = pos_df, y = meta_df, by = 'cell_ID')[c('cell_ID', 'sdimx', 'sdimy', 'Cell_Type', 'list_ID')]  # change the Cell_Type to your own cell type annotation
colnames(output_df) = c('Cell_ID', 'x', 'y', 'Cell_Type', 'Sample')
write.csv(output_df, file='ONTraC_input.csv', quote=FALSE, row.names=FALSE)
```



#### Running ONTraC
```{bash}
source ~/.bash_profile
conda activate ONTraC
ONTraC -d ONTraC_input.csv --preprocessing-dir preprocessing_dir --GNN-dir GNN_dir --NTScore-dir NTScore_dir --device cuda     --epochs 1000 -s 42 --patience 100 --min-delta 0.001 --min-epochs 50 --lr 0.03     --hidden-feats 4 -k 6 --modularity-loss-weight 0.3 --regularization-loss-weight 0.1     --purity-loss-weight 300 --beta 0.03 2>&1 | tee merfish_subset.log

ONTraC_analysis -o analysis_output/merfish_subset -l merfish_subset.log -d ONTraC_input.csv --preprocessing-dir preprocessing_dir --GNN-dir GNN_dir --NTScore-dir NTScore_dir -r -s
```

### visualization

- niche cluster
- NT score
- downstream trajectory analysis (Georage should present this topic on the morning session)

###
```{r}
NTScore = read.csv('NTScore_dir/merfish_subset_NTScore/NTScore.csv.gz')[c('Cell_NTScore')]
rownames(NTScore) = giotto_obj@cell_metadata$cell$rna$cell_ID
giotto_obj <- addCellMetadata(giotto_obj, new_metadata = NTScore)
```

