---
editor_options: 
  markdown: 
    wrap: 72
---

# Contributing to Giotto

Jiaji George Chen

August 7th 2024

## Contribution guideline

To be updated...
<https://drieslab.github.io/Giotto_website/CONTRIBUTING.html>

We welcome contributions or suggestions from other developers. Please
contact us if you have questions or would like to discuss an addition or
major modifications to the Giotto main code. The source code for Giotto
Suite may be found on our GitHub repository.

## Coding Style

Following a particular programming style will help programmers read and
understand source code conforming to the style, and help to avoid
introducing errors. Here we present a small list of guidelines on what
is considered a good practice when writing R code in Giotto package.
Most of them are adapted from Bioconductor - coding style or Google’s R
Style Guide. These guidelines are preferences and strongly encouraged!

-   **Overall style**

    -   We follow the BioConductor styling. You can set this up easily
        by installing *biocthis* and *styler.*

        ```{r, eval=FALSE}
        # package installations
        BiocManager::install("biocthis")
        install.packages("styler")

        # styling a file
        b_style <- biocthis::bioc_style()
        styler::style_file(path = "[???]", transformers = b_style)

        # styling the active package (may lead to lots of conflicts)
        # !! This should only be done be core devs with a lot of caution and forewarning !!
        styler::style_pkg(transformers = b_style)
        ```

    -   setting your default indent size to be 4 spaces instead of 2 is
        also recommended.

<!-- -->

-   **Function types**

    -   **exported** - Core functionality for users to directly use.
        These should have clear names and documentation

    -   **exported utility** - Secondary functionalities that are
        helpful to also have available, but are not directly related to
        data processing, analysis, and visualization. Examples are
        `dt_to_matrix()` or `wrap_msg()`

        -   another reason for this type of function is because Giotto
            is modular and some functions that are not expected to be
            commonly used by end users also need to be exported so that
            they are available across the Giotto ecosystem.

    -   **internal** - Functions that are never intended to be used
        outside of a module package. These are functions only relevant
        to the internals of one package, for example `.detect_in_dir()`
        from *Giotto's* internals which is pretty nondescript and mainly
        there to help with code organization.

-   **Naming**

    -   Use `camelCase` for **exported** functions. ex: `functionName()`

    -   Use `snake_case` for **exported utiliity** functions. ex:
        `function_name()`

    -   Use `.` prefix AND `snake_case` for **internal** functions. ex:
        `.function_name()`

    -   Use `snake_case` for parameter/argument names.

    -   Never use `.` as a separator in function naming. (in the S3
        class system, `fun(x)` where `x` is class foo will dispatch to
        `fun.foo()`)

-   **Use of symbols** Do not use any non-UTF-8 characters unless
    provided as the escape code. For example: `\u00F6` for `ö` Beyond
    these guidelines, *styler* should be used in order to maintain code
    uniformity.

## Stat functions

Most Giotto commands can accept several matrix classes (`DelayedMatrix`,
`SparseM`, Matrix or base `matrix`). To facilitate this we provide
flexible wrappers that work on any type of matrix class.

-   `mean_flex()`: analogous to `mean()`

-   `rowSums_flex()`: analogous to `rowSums()`

-   `rowMeans_flex()`: analogous to `rowMeans()`

-   `colSums_flex()`: analogous to `colSums()`

-   `colMeans_flex()`: analogous to `colMeans()`

-   `t_flex()`: analogous to `t()`

-   `cor_flex()`: analogous to `cor()`

## Auxiliary functions

Giotto has a number of auxiliary or convenience functions that might
help you to adapt your code or write new code for Giotto. We encourage
you to use these small functions to maintain uniformity throughout the
code.

-   `lapply_flex()`: analogous to lapply() and works for both windows
    and unix systems

-   `all_plots_save_function()`: compatible with Giotto instructions and
    helps to automatically save generated plots

-   `plot_output_handler()`: further wraps all_plots_save_function and
    includes handling for return_plot and show_plot and Giotto
    instructions checking

-   `determine_cores()`: determine the number of cores to use if a user
    does not set this explicitly

-   `get_os()`: identify the operating system

-   `update_giotto_params()`: will catch and store the parameters for
    each used command on a `giotto` object

-   `wrap_txt()`, `wrap_msg()`, etc: text and message formatting
    functions

-   `vmsg()`: framework for Giotto’s verbosity-flagged messages

-   `package_check()`: to check if a package exists, works for packages
    on CRAN, Bioconductor and Github

    -   Should be used within your contribution code if it requires the
        use of packages not in *Giotto's* `DESCRIPTION` file's depends
        imports section.

    -   Has the additional benefit that it will suggest to the user how
        to download the package if it is not available. To keep the size
        of *Giotto* within limits we prefer not to add too many new
        dependencies.

## Package Imports

*Giotto* tracks packages and functions to import in a centralized
manner. When adding code that requires functions from another package,
add the *roxygen* tags to the `package_imports.R` file for that *Giotto*
module.

## Getters and Setters

*Giotto* stores information in different
[slots](https://drieslab.github.io/Giotto_website/articles/articles/structure.html#giotto-object-structure),
which can be accessed through these getters and setters functions. They
can be found in the
[`accessors.R`](https://github.com/drieslab/Giotto/blob/suite/R/accessors.R)
file.

`setGiotto()`: Sets any *Giotto* subobject

`getCellMetadata()`: Gets cell metadata

`setCellMetadata()`: Sets cell metadata

`getFeatureMetadata()`: Gets feature metadata

`getFeatureMetadata()`: Sets feature metadata

`getExpression()`: To select the expression matrix to use

`setExpression()`: Sets a new expression matrix to the expression slot

`getSpatialLocations()`: Get spatial locations to use

`setSpatialLocations()`: Sets new spatial locations

`getDimReduction()`: To select the dimension reduction values to use

`setDimReduction()`: Sets new dimension reduction object

`getNearestNetwork()`: To select the nearest neighbor network (kNN or
sNN) to use

`setNearestNetwork()`: Sets a new nearest neighbor network (kNN or sNN)

`getSpatialNetwork()`: To select the spatial network to use

`setSpatialNetwork()`: Sets a new spatial network

`getPolygonInfo()`: Gets spatial polygon information

`setPolygonInfo()`: Set new spatial polygon information

`getFeatureInfo()`: Gets spatial feature information

`setFeatureInfo()`: Sets new spatial feature information

`getSpatialEnrichment()`: Gets spatial enrichment information

`setSpatialEnrichment()`: Sets new spatial enrichment information

`getMultiomics()`: Gets multiomics information

`setMultiomics()`: Sets multiomics information

## Python code

To use Python code we prefer to create a python wrapper/functions around
the python code, which can then be sourced by _reticulate_. As an example
we show the basic principles of how we implemented the Leiden clustering
algorithm.

1.  write python wrapper and store as `python_leiden.py` in
    `/inst/python`:

```{python, eval=FALSE, python.reticulate = FALSE}
import igraph as ig 
import leidenalg as la 
import pandas as pd
import networkx as nx

def python_leiden(df, partition_type, initial_membership=None, weights=None, n_iterations=2, seed=None, resolution_parameter = 1):
    
    # create networkx object
    Gx = nx.from_pandas_edgelist(df = df, source = 'from', target =  'to', edge_attr = 'weight')  

    # get weight attribute
    myweights = nx.get_edge_attributes(Gx, 'weight')

    ....

    return(leiden_dfr)
```

2.  source python code with *reticulate*:

```{r, eval=FALSE}
python_leiden_function = system.file("python", "python_leiden.py", package = 'Giotto') reticulate::source_python(file = python_leiden_function)
```

3.  use python code as if R code: See `doLeidenCLuster()` for more
    detailed information.

```{python, eval=FALSE, python.reticulate = FALSE}
pyth_leid_result = python_leiden(
    df = network_edge_dt,
    partition_type = partition_type, 
    initial_membership = init_membership, 
    weights = 'weight', 
    n_iterations = n_iterations,
    seed = seed_number, 
    resolution_parameter = resolution
)
```

